Gaia's constraints on the $\sim$$150\,M_\odot$ runaway star candidate VFTS682

### 2018-06-08 VFTS682_APJL.tex
minor bibliography
better caption
new paragraph and figure

% Using the stars selected as described in \Secref{data:gaia} (a
% subset is shown in blue in \Figref{fig:main}), we find averaged proper motion components of
% $\langle\mu_\mathrm{RA}\rangle = 1.7240\pm0.0002\,\mathrm{mas\ yr^{-1}}$ and
% $\langle\mu_\mathrm{DEC}\rangle = 0.6986\pm0.0003\,\mathrm{mas\
% yr^{-1}}$. These values are
% in good agreement with what is found by \cite{lennon:18}. However, we
% emphasize their sensitivity to the choice of stars adopted to define
% the reference frame.
Subtracting the mean proper motion components given by
\citet{lennon:18} from the
proper motion of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
components of proper motion of the star relative to the surrounding region
$\delta\mu_\mathrm{RA}^{Gaia} = 0.10 \pm 0.07\,\mathrm{mas\ yr^{-1}}$
and $\delta\mu_\mathrm{DEC}^{Gaia} = 0.08
\pm 0.09\,\mathrm{mas\ yr^{-1}}$. These components result in a
two-dimensional relative proper motion of $\delta \mu^{Gaia}=0.13\pm 0.09\,\mathrm{mas\
  yr^{-1}}$.
% \begin{equation}
%   \label{eq:pm_gaia_around}
%   \delta \mu^{Gaia} = \sqrt{\left(\delta\mu_\mathrm{RA}^{Gaia}\right)^2+\left(\delta\mu_\mathrm{DEC}^{Gaia}\right)^2}
%   = 0.13\pm 0.09\,\mathrm{mas\
%   yr^{-1}} \ \ .
% \end{equation}

Figure \Figref{fig:pm_polar} shows the distribution in relative proper motion
of the stars used in \cite{lennon:18} to define the frame of reference
we also adopt here, cf.~the right panel in their Fig.~2. The
proper motions are decomposed in the radial and tangential direction
from R136, to highlight the likelihood of it being the origin of fast
moving stars. The plus sign marks VFTS682, which sits at the edge of
the proper motion distribution, but is not a clear outlier. Most
relevant is the direction of the proper motion of VFTS682, which we
discuss in \Secref{sec:r136_origin}.

\begin{figure}%[htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{figures/figure_polar_682-1.pdf}
  \caption{Polar plot of the relative proper motion components for the stars
    defining our reference frame. VFTS682 is indicated by the plus sign, and
    other notable outliers \citep[see][]{lennon:18} are also labeled. Concentric dashed-lines denote relative proper motions of
    0.1, 0.2, and 0.4$\,\mathrm{mas\ yr^{-1}}$. Positive angle
    indicate a tangential component pointing in the counterclockwise
    direction, with 0 degrees corresponding to proper motion pointing
    radially outward from R136. The proper motion for VFTS682
    corresponds to almost radial motion away from the cluster.
  }
  \label{fig:pm_polar}
\end{figure}

@@

@@

Although the large uncertainties on the relative proper motion
components results in a wide range of possible directions,we argue that the most likely origin of the star is R136,
as suggested by the N-body simulations of \cite{fujii:11}
and \cite{banerjee:12}.

@@

\begin{figure}%[tbp]
  \centering
  \includegraphics[width=0.48\textwidth]{./figures/main_plot_good_gaia_only}  
  \caption{The red solid arrow indicates the proper motion of VFTS682
    relative to the region from \emph{Gaia} DR2, starting from the present day position of
    the star. The yellow arrows indicate the possible
    directions of projected motion within the \emph{Gaia} DR2 errors, and are extended
    backwards (dashed) to illustrate the uncertainty on the origin of the
    star. The length of the prolongations is proportional to the relative proper motion
    times the kinetic age of VFTS682.
    % \todo{load color picture on background? Fig with HST data
    % available, but confusing...}
  }

>>>>

% Using the stars selected as described in \Secref{data:gaia} (a
% subset is shown in blue in \Figref{fig:main}), we find averaged proper motion components of
% $\langle\mu_\mathrm{RA}\rangle = 1.7240\pm0.0002\,\mathrm{mas\ yr^{-1}}$ and
% $\langle\mu_\mathrm{DEC}\rangle = 0.6986\pm0.0003\,\mathrm{mas\
% yr^{-1}}$. These values are
% in good agreement with what is found by \cite{lennon:18}. However, we
% emphasize their sensitivity to the choice of stars adopted to define
% the reference frame.
Subtracting the mean proper motion components given by
\citet{lennon:18} from the
proper motion of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
components of proper motion of the star relative to the surrounding region
$\delta\mu_\mathrm{RA}^{Gaia} = 0.10 \pm 0.07\,\mathrm{mas\ yr^{-1}}$
and $\delta\mu_\mathrm{DEC}^{Gaia} = 0.08
\pm 0.09\,\mathrm{mas\ yr^{-1}}$. These components result in a
two-dimensional relative proper motion of $\delta \mu^{Gaia}=0.13\pm 0.09\,\mathrm{mas\
  yr^{-1}}$.

% \begin{equation}
%   \label{eq:pm_gaia_around}
%   \delta \mu^{Gaia} = \sqrt{\left(\delta\mu_\mathrm{RA}^{Gaia}\right)^2+\left(\delta\mu_\mathrm{DEC}^{Gaia}\right)^2}
%   = 0.13\pm 0.09\,\mathrm{mas\
%   yr^{-1}} \ \ .
% \end{equation}

% Figure \Figref{fig:pm_polar} shows the distribution in relative proper motion
% of the stars used in \cite{lennon:18} to define the frame of reference
% we also adopt here, cf.~the right panel in their Fig.~2. The
% proper motions are decomposed in the radial and tangential direction
% from R136, to highlight the likelihood of it being the origin of fast
% moving stars. The plus sign marks VFTS682, which sits at the edge of
% the proper motion distribution, but is not a clear outlier. Most
% relevant is the direction of the proper motion of VFTS682, which we
% discuss in \Secref{sec:r136_origin}.

% \begin{figure}%[htbp]
%   \centering
%   \includegraphics[width=0.5\textwidth]{figures/figure_polar_682-1.pdf}
%   \caption{Polar plot of the relative proper motion components for the stars
%     defining our reference frame. VFTS682 is indicated by the plus sign, and
%     other notable outliers \citep[see][]{lennon:18} are also labeled. Concentric dashed-lines denote relative proper motions of
%     0.1, 0.2, and 0.4$\,\mathrm{mas\ yr^{-1}}$. Positive angle
%     indicate a tangential component pointing in the counterclockwise
%     direction, with 0 degrees corresponding to proper motion pointing
%     radially outward from R136. The proper motion for VFTS682
%     corresponds to almost radial motion away from the cluster.
%   }
%   \label{fig:pm_polar}
% \end{figure}

@@

Figure~\ref{fig:mu_dist} shows the distribution in proper motion relative to R136
of OB type and Wolf-Rayet stars included both in the VFTS survey and
\emph{Gaia DR2} with reliable astrometric solutions. VFTS682 is not an
extreme outlier compared to the two O-type runaways (re-)identified by
\cite{lennon:18}. VFTS682 was expected to be a ``slow runaway'' by \citet{bestenlehner:11} in
the dynamical ejection scenario. The green shade in
\Figref{fig:mu_dist} shows the range of relative proper motions within
the uncertainties on the apparent age of the star, assuming it was
ejected from R136 very early in its life. The proper motion
measurements from \emph{Gaia} DR2 are consistent with the hypothesis
of dynamical ejection.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{figures/dist_mu_region.pdf}
  \caption{Distribution of OB-type and Wolf-Rayet stars in proper
    motion relative to R136. VFTS682 is not an extreme outlier, but
    its relative proper motion matches the expected value if it were indeed
    ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. The dark blue histogram contains 317 
    stars with error smaller than $0.1\,\mathrm{mas \
      yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc, the
    lighter blue histogram contains 36 stars with errors smaller than $0.05\,\mathrm{mas \
      yr^{-1}}$.}
  \label{fig:mu_dist}
\end{figure}

@@

Although the large uncertainties on the relative proper motion
components results in a wide range of possible directions,we argue that the most likely origin of the star is R136.

@@

\begin{figure}%[tbp]
  \centering
  \includegraphics[width=0.48\textwidth]{./figures/main_plot_good_gaia_only}  
  \caption{The red solid arrow indicates the proper motion of VFTS682
    relative to the region from \emph{Gaia} DR2, starting from the present day position of
    the star. The yellow arrows indicate the possible
    directions of projected motion within the \emph{Gaia} DR2 errors, and are extended
    backwards (dashed) to illustrate the uncertainty on the origin of the
    star. The length of the prolongations is proportional to the relative proper motion
    times the age of VFTS682 \citep[$1.0\pm0.2$\,Myr,][]{schneider:18}.
    % \todo{load color picture on background? Fig with HST data
    % available, but confusing...}
  }
